\acknowledgment

A page or two so of shout-outs to people you appreciate.  Don't forget
your advisor and committee members!

I would like to thank everyone in my life for whom writing this dissertation was made possible.
Suppose the cross-section of life-changing interactions across many spacetime scales is given by
\begin{equation}
\label{eqn:clayton-cross-section}
d\sigma^{\text{Clayton}} = d\sigma^{\text{childhood}} \otimes d\sigma^{\text{college}} \otimes d\sigma^{\text{gradschool}}
\end{equation}  
The first term in \ref{eqn:clayton-cross-section} cannot exist without the love and support of my mother and father, Robin Schroeder and Randall Bennett, respectively, along with the loving support of my stepmother, Lisa Bennett.
This term also contains import contributions from my closest childhood friends, Sam Hemes and Connor Eccles.
Together, we made movies, we made music, and most importantly, we made memories.
The second term in \ref{eqn:clayton-cross-section} cannot be expressed accuratately without Emily Hegland -- the love of my life -- for whom I have been dedicated to since the day the we met.
This term also has non-negligable contributions from Tyson Williams, for whom inspired me first to study physics.
Finally, the last, and perhaps most important term in application to this thesis, derived from my time at UIC, immersing myself in the deep and beautiful laws of the universe:  I must first thank Olga Evdokimov, my advisor.
From you I learned so much, most of which is the 





\initials{CMB}
